---
title: "Bioinformatics_Analyisis"
author: "Jorlen Moreno"
date: "2026-04-27"
output: html_document
---

Load the library
```{r}
library(seqinr)

```

Read the fasta files
```{r}
austria_data <- read.fasta("austria.fasta")
brazil_data <- read.fasta("brazil.fasta")
india_data <- read.fasta("india.fasta")

seq_austria <- austria_data[[1]]
seq_brazil  <- brazil_data[[1]]
seq_india   <- india_data[[1]]
```


We make a list out of the datasets
```{r}
seqs_list <- list(
  Austria = seq_austria,
  Brazil  = seq_brazil,
  India   = seq_india
)
seqs_list
```

Then, we will calculate the length of each sequence and will graph it.
```{r}
lengths <- sapply(seqs_list, length)

print("Sequence longitudes")
print(lengths)

#we create the plot for each one
barplot(lengths,
        main = "Viral sequences lengths",
        col = hcl.colors(4, "Plasma"),
        ylab = "Nucleotides",
        xlab = "Origin$")

```
Then we visualize nucleotide composition.
```{r}
comp_nucleotides <- sapply(seqs_list, function(x) count(x, 1))

# Barplot to count ACGT nucleotides
barplot(comp_nucleotides,
        beside = TRUE, 
        main = "Nucleotide composition by origin",
        col = hcl.colors(4, "Plasma"),
        legend = rownames(comp_nucleotides),
        xlab = "Origin",
        ylab = "Frequence")
```
Get the GC percentage
```{r}

gc_percentage <- sapply(seqs_list, GC) * 100

print("GC content:")
print(gc_percentage)

#gc barplot
barplot(gc_percentage,
        main = "GC content (%)",
        col = hcl.colors(4, "Plasma"),
        ylab = "% GC",
        xlab = "Origin",
        ylim = c(0, 60)) 

```
Finally, we get the complimentray sequence

```{r}

comp_austria <- c2s(comp(seq_austria))
comp_brazil  <- c2s(comp(seq_brazil))
comp_india   <- c2s(comp(seq_india))

print("complementary sequences:")
print(substr(comp_austria, 1, 50))
print(substr(comp_brazil, 1, 50))
print(substr(comp_india, 1, 50))


